---
title: "In-Class Exercise 4"
author: "Yang Jun"
date: 6 May 2023
date-modified: "`r Sys.Date()`"
execute:
  echo: true
  eval: true
  warning: false
---

## Import Packages and Data

```{r}
pacman::p_load(rstatix, gt, patchwork, tidyverse)
```

```{r}
exam_data <- read_csv('data/Exam_data.csv')
```

## Visualising a Normal Distribution

### First Attempt: QQ Plot

We can use a QQ plot to visualise whether a distribution is normal or not. In the plot below, the points deviate significantly from the straight line, indicating that the data is *not* normally distributed.

::: panel-tabset
## The Plot

```{r}
#| echo: false
ggplot(data = exam_data,
       aes(sample=ENGLISH)) +
  stat_qq() +
  stat_qq_line()
```

## The Code Chunk
```{r}
#| eval: false
ggplot(data = exam_data,
       aes(sample=ENGLISH)) +
  stat_qq() +
  stat_qq_line()
```

::: callout-note
## Note

We use `stat_qq()` and `stat_qq_line()` methods to plot the QQ plot. Note that here `aes` takes an argument called `sample` instead of typical `x` and/or `y`.
:::

:::

### Second Attempt: QQ Plot + Statistical Test Table

We can add a table showing the results of a formal statistical test for normality. Here we use the Shapiro-Wilk Test.

::: panel-tabset
## The Plot

```{r}
#| echo: false
qq <- ggplot(data = exam_data,
             aes(sample=ENGLISH)) +
  stat_qq() +
  stat_qq_line()

sw_t <- exam_data %>%
  shapiro_test(ENGLISH) %>%
  gt()

tmp <- tempfile(fileext = '.png')
gtsave(sw_t, tmp)
table_png <- png::readPNG(tmp, native=TRUE)

qq + table_png
```

## The Code Chunk
```{r}
#| eval: false
qq <- ggplot(data = exam_data,
             aes(sample=ENGLISH)) +
  stat_qq() +
  stat_qq_line()

sw_t <- exam_data %>%
  shapiro_test(ENGLISH) %>%
  gt()

tmp <- tempfile(fileext = '.png')
gtsave(sw_t, tmp)
table_png <- png::readPNG(tmp, native=TRUE)

qq + table_png
```

:::

