---
title: "Hands-On Exercise 1"
author: "Yang Jun"
---

## Getting Started

### Install and Load R Packages

The code chunk below uses `p_load()` of `pacman` package to check if `tidyverse` packages are installed in the computer. If they are, they will be loaded into R.

```{r}
pacman::p_load(tidyverse)
```

### Import the Data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

## Plotting Charts

### Bar Chart

```{r}
ggplot(data=exam_data,
       aes(x=RACE)) +
  geom_bar()
```

### Dot Plot

```{r}
ggplot(data=exam_data,
       aes(x=MATHS)) +
  geom_dotplot(binwidth=2.5,
               dotsize=0.5) +
  scale_y_continuous(NULL,
                     breaks=NULL)
```

### Histogram 1

```{r}
ggplot(data=exam_data,
       aes(x=MATHS)) +
  geom_histogram(bins=20,
                 color='black',
                 fill='light blue')
```

### Histogram 2

```{r}
ggplot(data=exam_data,
       aes(x=MATHS,
           fill=GENDER)) +
  geom_histogram(bins=20,
                 color='white')
```

### Kernel Density Estimate

```{r}
ggplot(data=exam_data, 
       aes(x=MATHS,
           colour=GENDER)) +
  geom_density()
```

### Boxplot

```{r}
ggplot(data=exam_data,
       aes(y=MATHS,
           x=GENDER)) +
  geom_boxplot(notch=TRUE)
```

### Violin Plot

Useful to compare several distributions.

```{r}
ggplot(data=exam_data,
       aes(y=MATHS,
           x=GENDER)) +
  geom_violin()
```

### Scatterplot

```{r}
ggplot(data=exam_data,
       aes(x=MATHS,
           y=ENGLISH)) +
  geom_point()
```

### Combining Geom Types

```{r}
ggplot(data=exam_data,
       aes(y=MATHS,
           x=GENDER)) +
  geom_boxplot() +
  geom_point(position='jitter',
             size=0.5)
```

## Adding Statistics

### Adding mean to boxplot

```{r}
ggplot(data=exam_data,
       aes(y=MATHS,
           x=GENDER)) +
  geom_boxplot() +
  stat_summary(geom='point',
               fun='mean',
               colour='red',
               size=4)
```

### Adding best fit curve to scatterplot

```{r}
ggplot(data=exam_data,
       aes(x=MATHS,
       y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm,
              size=0.5)
```

## Facets

### Facet Wrap

```{r}
ggplot(data=exam_data,
       aes(x=MATHS)) +
  geom_histogram(bins=20) +
  facet_wrap(~CLASS)
```

### Facet Grid

```{r}
ggplot(data=exam_data,
       aes(x=MATHS,
           y=ENGLISH)) +
  geom_point() +
  facet_grid(~CLASS)
```

## Coordinates

### Flipping vertical bar chart to horizontal

```{r}
ggplot(data=exam_data,
       aes(x=RACE)) +
  geom_bar() +
  coord_flip()
```

### Changing axis ranges

```{r}
ggplot(data=exam_data,
       aes(x=MATHS,
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm,
              size=0.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
```

## Themes

```{r}
ggplot(data=exam_data,
       aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_minimal()
```
