---
title: "Hands-On Exercise 4"
author: "Yang Jun"
date: 4 May 2023
---

In this Hands-On Exercise, we explore statistical analysis and uncertainty using visual analytics.

# Visual Statistical Analysis

## Getting Started

### Install and Load Required Packages

```{r}
pacman::p_load(ggstatsplot, tidyverse)
```

### Import the Data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

## One-sample test

The following code uses the `gghistostats()` method to build a visual of a one-sample test for English scores.

```{r}
set.seed(1234)

gghistostats(
  data = exam_data,
  x = ENGLISH,
  type = 'bayes',
  test.value = 60,
  xlab = 'English scores'
)
```

## Two-sample means test

The following code uses the `ggbetweenstats()` method to build a visual for a two-sample means test of Maths scores by gender.

```{r}
ggbetweenstats(
  data = exam_data,
  x = GENDER,
  y = MATHS,
  type = 'np',
  messages = FALSE
)
```

## One-way ANOVA test

The following code uses the `ggbetweenstats()` method to build a visual for an ANOVA test of English scores by race.

```{r}
ggbetweenstats(
  data = exam_data,
  x = RACE,
  y = ENGLISH,
  mean.ci = TRUE,
  pairwise.comparisons = TRUE,
  pairwise.display = 's',
  p.adjust.method = 'fdr',
  messages = FALSE
)
```

## Test for significance of correlation

The following code uses the `ggscatterstats()` method to build a visual for testing the significance of correlation between Maths and English scores.

```{r}
ggscatterstats(
  data = exam_data,
  x = MATHS,
  y = ENGLISH,
  marginal = FALSE
)
```

## Test for significance of association

The following code uses the `ggbarstats()` method to build a visual for testing the significance of association between binned Maths scores and gender.

```{r}
exam_binned <- exam_data %>%
  mutate(MATHS_binned =
           cut(MATHS, breaks = c(0,60,75,85,100)))
```

```{r}
ggbarstats(
  data = exam_binned,
  x = MATHS_binned,
  y = GENDER
)
```

# Visualising Models

## Getting Started

### Install and Load Required Packages

```{r}
pacman::p_load(readxl, performance, parameters, see)
```

### Import the Data

```{r}
car_resale <- read_xls('data/ToyotaCorolla.xls', 'data')
car_resale
```

## Multiple Regression Model

### Create the Model

```{r}
model <- lm(Price ~ Age_08_04 + Mfg_Year + KM + 
              Weight + Guarantee_Period, data = car_resale)
model
```

### Check for Multicollinearity

```{r}
check_c <- check_collinearity(model)
plot(check_c)
```

### Check for Normality

```{r}
model1 <- lm(Price ~ Age_08_04 + KM + 
              Weight + Guarantee_Period, data = car_resale)
```

```{r}
check_n <- check_normality(model1)
plot(check_n)
```

### Check for homogeneity of variances

```{r}
check_h <- check_heteroscedasticity(model1)
plot(check_h)
```

### Complete check

```{r}
check_model(model1)
```

### Visualising regression parameters using `plot()`

```{r}
plot(parameters(model1))
```

### Visualising regression parameters using ggcoefstats()

```{r}
ggcoefstats(model1, output='plot')
```

# Visualising Uncertainty

## Getting Started

### Install and Load Required Packages

```{r}
pacman::p_load(tidyverse, plotly, crosstalk, DT, ggdist, gganimate)
```

### Import the Data

```{r}
exam <- read_csv("data/Exam_data.csv",
                 show_col_types = FALSE)
```

## Uncertainty of Point Estimates

Compute the count of observations, mean, standard deviation and standard error of Maths scores by RACE:

```{r}
my_sum <- exam %>%
  group_by(RACE) %>%
  summarise(
    n = n(),
    mean = mean(MATHS),
    sd = sd(MATHS)
  ) %>%
  mutate(
    se = sd/sqrt(n-1)
  )
```

```{r}
knitr::kable(head(my_sum),
             format = 'html')
```

### Standard error of mean Maths scores

```{r}
ggplot(my_sum) +
  geom_errorbar(
    aes(x=RACE,
        ymin=mean-se,
        ymax=mean+se),
    width = 0.2,
    color = 'black',
    alpha = 0.9,
    size = 0.5
  ) +
  geom_point(
    aes(x=RACE,
        y=mean),
    stat = 'identity',
    color = 'red',
    size = 1.5,
    alpha = 1
  ) +
  ggtitle("Standard error of mean Maths scores by Race")
```

### 95% CI of mean Maths scores

```{r}
ggplot(my_sum) +
  geom_errorbar(
    aes(x=reorder(RACE,-mean),
        ymin=mean-se*qnorm(1-0.025),
        ymax=mean+se*qnorm(1-0.025)),
    width = 0.2,
    color = 'black',
    alpha = 0.9,
    size = 0.5
  ) +
  geom_point(
    aes(x=reorder(RACE,-mean),
        y=mean),
    stat = 'identity',
    color = 'red',
    size = 1.5,
    alpha = 1
  ) +
  ggtitle("95% confidence interval for mean Maths scores by Race")
```

### Interactive error bars

```{r}

```

## Visualising uncertainty using `ggdist` methods

### Using `stat_pointinterval()`

```{r}
exam %>%
  ggplot(aes(x = RACE,
             y = MATHS)) +
  stat_pointinterval() +
  labs(
    title = "Visualising confidence intervals for mean Maths scores",
    subtitle = "Mean Point + Multiple-interval plot"
  )
```

### Using `stat_gradientinterval()`

```{r}
exam %>%
  ggplot(aes(x = RACE,
             y = MATHS)) +
  stat_gradientinterval(
    fill = 'skyblue',
    show.legend = TRUE
  ) +
  labs(
    title = "Visualising confidence intervals for mean Maths scores",
    subtitle = "Gradient + interval plot"
  )
```

### Using Hypothetical Outcome Plots (HOPs)

```{r}
library(ungeviz)
```

```{r}
ggplot(data = exam, 
       (aes(x = factor(RACE), y = MATHS))) +
  geom_point(position = position_jitter(
    height = 0.3, width = 0.05), 
    size = 0.4, color = "#0072B2", alpha = 1/2) +
  geom_hpline(data = sampler(25, group = RACE), height = 0.6, color = "#D55E00") +
  theme_bw() + 
  # `.draw` is a generated column indicating the sample draw
  transition_states(.draw, 1, 3)
```

# Funnel Plots

## Getting Started

### Install and Load Required Packages

```{r}
pacman::p_load(tidyverse, FunnelPlotR, plotly, knitr)
```

### Import the Data

```{r}
covid19 <- read_csv('data/COVID-19_DKI_Jakarta.csv', show_col_types = FALSE) %>%
  mutate_if(is.character, as.factor)
```

## Basic Funnel Plot using FunnelPlotR

```{r}
funnel_plot(
  numerator = covid19$Death,
  denominator = covid19$Positive,
  group = covid19$`Sub-district`
)
```

### Funnel Plot: Makeover 1

```{r}
funnel_plot(
  numerator = covid19$Death,
  denominator = covid19$Positive,
  group = covid19$`Sub-district`,
  data_type = "PR",
  x_range = c(0, 6500),
  y_range = c(0, 0.05)
)
```

### Funnel Plot: Makeover 2

```{r}
funnel_plot(
  numerator = covid19$Death,
  denominator = covid19$Positive,
  group = covid19$`Sub-district`,
  data_type = "PR",
  x_range = c(0, 6500),
  y_range = c(0, 0.05),
  label = NA,
  title = "Cumulative COVID-19 Fatality Rate by Cumulative Total Number of COVID-19 Positive Cases",
  x_label = "Cumulative COVID-19 Positive Cases",
  y_label = "Cumulative Fatality Rate"
)
```

## Building Funnel Plots from Scratch

### Compute derived fields

```{r}
df <- covid19 %>%
  mutate(rate = Death/Positive) %>%
  mutate(rate.se = sqrt( (rate*(1-rate))/Positive )) %>%
  filter(rate > 0)
```

```{r}
fit.mean <- weighted.mean(df$rate, 1/df$rate.se^2)
```

### Calculate lower and upper limits for 95% and 99.9% CI

```{r}
number.seq <- seq(1, max(df$Positive), 1)
number.ll95 <- fit.mean - 1.96 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ul95 <- fit.mean + 1.96 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ll999 <- fit.mean - 3.29 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ul999 <- fit.mean + 3.29 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
dfCI <- data.frame(number.ll95, number.ul95, number.ll999, 
                   number.ul999, number.seq, fit.mean)
```

### Plot static funnel plot

```{r}
p <- ggplot(data=df,
            aes(x=Positive,
                y=rate
                )) +
  geom_point(aes(label=`Sub-district`),
             alpha=0.4) +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ll95), 
            size = 0.4, 
            colour = "grey40", 
            linetype = "dashed") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ul95), 
            size = 0.4, 
            colour = "grey40", 
            linetype = "dashed") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ll999), 
            size = 0.4, 
            colour = "grey40") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ul999), 
            size = 0.4, 
            colour = "grey40") +
  geom_hline(data = dfCI, 
             aes(yintercept = fit.mean), 
             size = 0.4, 
             colour = "grey40") +
  coord_cartesian(ylim=c(0,0.05)) +
  annotate("text", x = 200, y = 0.003, label = "95%", size = 3, colour = "grey40") +
  annotate("text", x = 500, y = 0.0, label = "99%", size = 3, colour = "grey40") + 
  ggtitle("Cumulative Fatality Rate by Cumulative Number of COVID-19 Cases") +
  xlab("Cumulative Number of COVID-19 Cases") + 
  ylab("Cumulative Fatality Rate") +
  theme_light() +
  theme(plot.title = element_text(size=12),
        legend.position = c(0.91,0.85), 
        legend.title = element_text(size=7),
        legend.text = element_text(size=7),
        legend.background = element_rect(colour = "grey60", linetype = "dotted"),
        legend.key.height = unit(0.3, "cm"))


p
```

### Adding interactivity

```{r}
fp_ggplotly <- ggplotly(p,
                        tooltip = c('label', 'x', 'y'))

fp_ggplotly
```
