---
title: "Hands-On Exercise 2"
author: "Yang Jun"
---

## Getting Started

### Install and Load Required Packages

The code chunk below will be used to check if these packages have been installed and also will load them onto your working R environment:

```{r}
pacman::p_load(ggrepel, patchwork, ggthemes, hrbrthemes, tidyverse) 
```

### Import the Data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

## Annotations without overlapping labels

### Annotation using default ggplot `geom_label()`

```{r}
ggplot(data=exam_data,
       aes(x=MATHS,
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm,
              linewidth=0.5) +
  geom_label(aes(label=ID)) +
  coord_cartesian(xlim=c(0,100),ylim=c(0,100)) +
  ggtitle("English scores vs Maths scores for Primary 3 students")
```

### Annotation using ggrepel `geom_text_repel()`

```{r}
ggplot(data=exam_data,
       aes(x=MATHS,
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm,
              linewidth=0.5) +
  geom_label_repel(aes(label=ID)) +
  coord_cartesian(xlim=c(0,100),ylim=c(0,100)) +
  ggtitle("English scores vs Maths scores for Primary 3 students")
```

## Beyond ggplot2 Themes

### Using default ggplot2 `theme_gray` theme

```{r}
ggplot(data=exam_data,
       aes(x=MATHS)) +
  geom_histogram(bins=20,
                 color='grey75') +
  theme_gray() +
  ggtitle("Distribution of Maths scores")
```

### Using 'Economist' theme from `ggthemes`

```{r}
ggplot(data=exam_data,
       aes(x=MATHS)) +
  geom_histogram(bins=20,
                 color='grey75') +
  theme_economist() +
  ggtitle("Distribution of Maths scores")
```

### Using `hrbrthemes` package

```{r}
ggplot(data=exam_data,
       aes(x=MATHS)) +
  geom_histogram(bins=20,
                 color='grey75') +
  theme_ipsum() +
  ggtitle("Distribution of Maths scores")
```

## Plotting Multiple Graphs using `patchwork`

### Initial Graphs

```{r}
p1 <- ggplot(data=exam_data,
       aes(x=MATHS)) +
  geom_histogram(binwidth=5,
                 boundary=100,
                 color='grey75') +
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of Maths scores")

p1
```

```{r}
p2 <- ggplot(data=exam_data,
       aes(x=ENGLISH)) +
  geom_histogram(binwidth=5,
                 boundary=100,
                 color='grey75') +
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of English scores")

p2
```

```{r}
p3 <- ggplot(data=exam_data,
       aes(x=MATHS,
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm,
              linewidth=0.5) +
  coord_cartesian(xlim=c(0,100),ylim=c(0,100)) +
  ggtitle("English scores vs Maths scores for Primary 3 students")

p3
```

### Two Graphs

```{r}
p1 + p2
```

### Three Graphs

```{r}
(p1 / p2) | p3 +
  plot_annotation(tag_levels='I')
```

### Inset another graph

```{r}
p3 + inset_element(p2, 
                   left = 0.02, 
                   bottom = 0.7, 
                   right = 0.5, 
                   top = 1)
```

### Combining with themes

```{r}
patchwork <- (p1 / p2) | p3
patchwork & theme_economist()
```
